---
title: "Evaluate Node Coverage in bacterial WGA"
output: html_document
author: Nicolò Rossi
runtime: shiny
---

```{r, echo=F, results=F, warning=FALSE, include=F}
knitr::opts_chunk$set(echo = F, fig.align="center")
library(factoextra)
library(tidyverse)
library(mclust)
library(DT)
library(ggraph)
library(neighbr)
library(plotly)

### reactive variables declaration

options(shiny.maxRequestSize=1024^3)
rv <- reactiveValues()

rv$dataset <- data.frame()
rv$text <- ""
# plots
rv$Kmeans1D <- NULL
rv$Kmeans2D <- NULL
rv$mixture <- NULL
rv$custom <- NULL
rv$set_dataset <- FALSE

# load .summary dataset
load_dataset <- function(){
  try({
    filepath <- input$uploadedFile$datapath
    mean.report <- read_tsv(filepath) %>% 
      select(-X1) %>% # remove index
      arrange(desc(Coverage)) %>% 
      mutate(lcov=scale(log(Coverage)),llen=scale(log(Length))) # TODO
    rv$dataset <- mean.report
    rv$set_dataset <- TRUE
  })
}

ALPHA_VALUE <- 0.7
```


With this simple web application it is possible to assess the coverage distribution among the different nodes created by a Whole Genome Assembly pipeline, such as **SPAdes**.


### File upload 

Upload you coverage file. This must the `depth_summary` file created by the `check_coverage.py` script (`wgalp check-coverage` command):

```{r}
fileInput("uploadedFile", "Choose File", multiple = FALSE) %>% shiny::tagAppendAttributes(style = 'width: 100%;')
actionButton("loadData", "Load new dataset") %>% shiny::tagAppendAttributes(style = 'width: 100%;')
observeEvent(input$loadData, { load_dataset() })
```

### Tabular view

You can sort and filter the table as you prefer:

```{r}
renderDataTable({
  datatable(rv$dataset, options = list("pageLength" = 10), filter="bottom")
})
```

### Node distribution

```{r}

scatter_plotter <- function(p){ 
  renderPlotly({
    ggplotly(p, tooltip=c("x", "y", "text")) %>%
      layout(legend=list(title=list(text='<b> Clusters </b>'))) 
  }) 
}

# aux function for 1D k-means, computes the position of a line dividing the two clusters
get_boundary <- function(df, scaled=T){
    df1 <- df %>% filter(clusters==1)
    df2 <- df %>% filter(clusters==2)
    m1 <- mean(pull(df1,Coverage))
    m2 <- mean(pull(df2,Coverage))
    
    if(!scaled){
        if(m1 > m2){
            return( (min(pull(df1,Coverage)) + max(pull(df2,Coverage)))/2 )    
        }else{
            return( (max(pull(df1,Coverage)) + min(pull(df2,Coverage)))/2 )     
        }
    }else{
        if(m1 > m2){
            return( (min(pull(df1,lcov)) + max(pull(df2,lcov)))/2 )    
        }else{
            return( (max(pull(df1,lcov)) + min(pull(df2,lcov)))/2 )     
        }
    }
}

# 1 dimensional K means
Kmeans1D <- function(after=FALSE){
  out_plots <- list()
  if(after){
    mean.report <- rv$dataset %>% filter(input$boundary<Coverage)
  }else{
    mean.report <- rv$dataset
  }
  out_plots[[1]] <- fviz_nbclust(mean.report %>% select(lcov), kmeans, method = "silhouette")
  kmeans_clusters1D <- kmeans(mean.report %>% select(lcov),2)
  mean.report <- mean.report %>% mutate(clusters = kmeans_clusters1D$cluster)
  out_plots[[2]] <- ggplot(mean.report) + 
      geom_point(aes(x=Coverage,y=Length, text=mean.report$Name, color=as.factor(clusters)), alpha=ALPHA_VALUE) +
      geom_vline(xintercept = get_boundary(mean.report, scaled=F) ) +
      scale_x_log10() + 
      scale_y_log10()
  
  if(after){
    rv$Kmeans1D_after <- scatter_plotter(out_plots[[2]])
  }else{
    rv$Kmeans1D <- scatter_plotter(out_plots[[2]])  
  }
  
}

# 2 dimensional K means
Kmeans2D <- function(after=FALSE){
  out_plots <- list()
  if(after){
    mean.report <- rv$dataset %>% filter(input$boundary<Coverage)
  }else{
    mean.report <- rv$dataset
  }
  out_plots[[1]] <- fviz_nbclust(mean.report %>% select(lcov, llen), kmeans, method = "silhouette")
  kmeans_clusters1D <- kmeans(mean.report %>% select(lcov, llen),4)
  mean.report <- mean.report %>% mutate(clusters = kmeans_clusters1D$cluster)
  out_plots[[2]] <- ggplot(mean.report) + 
    geom_point(aes(x=Coverage,y=Length, text=mean.report$Name, color=as.factor(clusters)), alpha=ALPHA_VALUE) +
    scale_x_log10() + 
    scale_y_log10() 
  
  if(after){
    rv$Kmeans2D_after <- scatter_plotter(out_plots[[2]])  
  }else{
    rv$Kmeans2D <- scatter_plotter(out_plots[[2]])
  }
  
}

# gaussian misture model
mclust_model <- function(after=FALSE){
  # mixture clustering
  if(after){
    mean.report <- rv$dataset %>% filter(input$boundary<Coverage)
  }else{
    mean.report <- rv$dataset
  }
  BIC <- mclustBIC(mean.report %>% select(lcov, llen))
  mclust_model <- mclustModel(mean.report %>% select(lcov, llen), BIC)
  mclust_clusters <- apply(mclust_model$z, 1, which.max)
  mean.report <- mean.report %>% mutate(clusters = mclust_clusters) 
  plot <- ggplot(mean.report) + 
    geom_point(aes(x=Coverage, y=Length, text=mean.report$Name, color=as.factor(clusters)), alpha=ALPHA_VALUE) +
    scale_x_log10() + 
    scale_y_log10() 
  
  if(after){
    rv$mixture_after <- scatter_plotter(plot)
  }else{
    rv$mixture <- scatter_plotter(plot)  
  }
  
}

# custom boundary
custom <- function(after = FALSE){
  limit <- input$boundary
  if(after){
    mean.report <- rv$dataset %>% filter(input$boundary<Coverage)
  }else{
    mean.report <- rv$dataset
  }
  plot <- ggplot(mean.report) + 
    geom_point(aes(x=Coverage, y=Length, text=mean.report$Name, color=limit<Coverage), alpha=ALPHA_VALUE) +
    geom_vline(xintercept = limit ) +
    scale_x_log10() + 
    scale_y_log10() 
  
  if(!after){
    rv$custom <- scatter_plotter(plot)
  }
  
  coverage_ <- ggplot(mean.report) + 
    geom_density(aes(x=Coverage), color="blue") +
    scale_x_log10()
  length_ <- ggplot(mean.report) +
    geom_density(aes(x=Length), color="red") +
    scale_x_log10()
  normsd_ <- ggplot(mean.report) +
    geom_density(aes(x=Sd/Coverage)) +
    labs(x="standardized (sd / coverage)")
  
  if(after){
    rv$coverage_after <- coverage_
    rv$length_after <- length_
    rv$normsd_after <- normsd_
  }else{
    rv$coverage <- coverage_
    rv$length <- length_
    rv$normsd <- normsd_
  }
    
}

# --- UI

numericInput("boundary", "Set custom boundary", 20, min=0)

renderUI({
  mainPanel(
      tabsetPanel(id = "inputTabs",
          tabPanel("custom", rv$custom),
          tabPanel("1D-Kmeans", rv$Kmeans1D),
          tabPanel("2D-Kmeans", rv$Kmeans2D),
          tabPanel("Mixture", rv$mixture),
          tabPanel("Coverage", renderPlot({rv$coverage})),
          tabPanel("Length", renderPlot({rv$length })),
          tabPanel("Normalized sd", renderPlot({rv$normsd}))
      )
  ) %>% shiny::tagAppendAttributes(style = 'width: 100%;')
})
  
reactive({
  if(rv$set_dataset){
    Kmeans1D()
    Kmeans2D()
    mclust_model()
    rv$set_dataset <- FALSE
  }
  custom()
})

```

### Selected nodes

```{r}
renderPrint({
  rv$dataset %>% filter(Coverage>=input$boundary) %>% pull(Name) %>% paste(collapse="\n") %>% cat
})
```

Copy these lines into a file to filter the nodes with `filter_fasta.py` script.

### After selection

Check the status after the application of the custom boundary. This can be useful to 
refine the analysis.

```{r}
renderUI({
  mainPanel(
      tabsetPanel(id = "inputTabs_after",
          tabPanel("1D-Kmeans", rv$Kmeans1D_after),
          tabPanel("2D-Kmeans", rv$Kmeans2D_after),
          tabPanel("Mixture", rv$mixture_after),
          tabPanel("Coverage", renderPlot({rv$coverage_after})),
          tabPanel("Length", renderPlot({rv$length_after})),
          tabPanel("Normalized sd", renderPlot({rv$normsd_after}))
      )
  ) %>% shiny::tagAppendAttributes(style = 'width: 100%;')
})

reactive({
  Kmeans1D(after=T)
  Kmeans2D(after=T)
  mclust_model(after=T)
  custom(after=T)
})

```









